### 2025-02-07 mixed-models.qmd
20241229 - bootstrap Cullen and Frey graph
20250207 - challenges of trajectories

@@

fitdistrplus::descdist(fantasyPointsVector)
```

>>>>

In this chapter, we use mixed models to demonstrate how to model trajectories (growth curves) of players' fantasy points as a function of age.
One of the challenges when modeling players' fantasy points as a function of age is that fantasy points are a function of both ability and opportunity.
With age, the player's ability and opportunity may both decline, but it is difficult to disentangle the extent to which a player's decline is related to declines in ability versus opportunity.
As players get older, they may be supplanted by younger, more talented players.
With age, despite still having latent (unobserved) ability, players will get fewer and, eventually, no opportunity.
And we do not know the counterfactual—how many fantasy points they would have scored had they been given the full opportunity each season.
Thus, players may go from scoring 100+ points in a season to all of a sudden scoring way fewer points, with little in the way of intermediate steps.
Thus, this should inform how you interpret the resulting trajectories.
For instance, the steep declines can make it look like the model-implied fantasy points go well below zero at later ages, even though it is obviously not likely for players to obtain such a magnitude of negative fantasy points.
These negative model-implied values of fantasy points at later ages is likely an artifact of the decreasing opportunity that players tend to get as they age.

@@

fitdistrplus::descdist(
  fantasyPointsVector,
  boot = 1000)
```
